### 2016-01-16 03-design.tex
Make figs 1x2 i.s.o. 2x1. Move them around

>>>>

% horrible hack to move the table up
\begin{table*}
  \centering
  \begin{tabular}{l r r r l}
    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
                    &            &               & communities    &             \\
    \hline
    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
    \hline
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{tab-datasets}
\end{table*}

### 2016-01-16 04-evaluation.tex
Typo fixes
Make figs 1x2 i.s.o. 2x1. Move them around

\begin{table*}
  \centering
  \begin{tabular}{l r r r l}
    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
                    &            &               & communities    &             \\
    \hline
    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
    \hline
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{tab-datasets}
\end{table*}

@@

\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\textwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure*}

In order to evaluate the horizontal scalability of the distributed
implementation we tested the system's performance across different cluster
sizes while holding the problem size constant. For this study, we used the
com-Friendster graph as it contains the largest number of vertices and edges.
%
Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
iterations across multiple cluster sizes. The x-axis starts from 8 worker nodes
as the data set is too large to fit into the collective memory of a smaller
cluster.
%
As shown in the figure, the execution
time steadily decreases by increasing the cluster size.
A deeper analysis of the individual execution phases of the algorithm provides
insights into the scalability of its building blocks.
%
In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
presents the cumulative time spent in individual computational phases across
iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
achieved for the same experiments reported in
Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
the execution is \textit{update\_phi\_pi}.
The reported total time for each cluster size is significantly less than the
sum of the execution times of the individual phases. This is due to the
overlapping execution of the two most expensive phases, namely,
\textit{update\_phi\_pi} and
minibatch deployment. Both of these phases initially gain significant speedup with the
addition of compute nodes. However, the speedup curve gradually slows down for
larger cluster sizes as the work granularity of each worker node decreases,
limiting their resource utilization.
%
The time spent in \textit{update\_beta\_theta} remains relatively constant across cluster sizes
as it performs an insignificant amount of work compared to the synchronization
overhead of a collective MPI reduction operation contained within it.

@@

\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\textwidth}
  \caption{(a) Average execution time per algorithm iteration varying
  the number of latent communities proportionally to the number of compute
  nodes. (b) The exact number of communities used for each data point in (a).}
  \label{fig-weak-scaling}
\end{figure*}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{l d{5.1} d{5.1}}
    \textbf{Iteration stages}  & \multicolumn{1}{c}{non-pipelined}
                                          & \multicolumn{1}{c}{pipelined} \\
    \hline
    total                      & 450    & 365 \\
    draw/deploy mini-batch      &  45.6  &      \\
    update\_phi                & 285    & 241 \\
    update\_pi                 &   3.8  &   4.6 \\
    update $\beta/\theta$      &  25.9  &   33.6 \\
    \hline
    \textbf{Substages within update\_phi} \\
    draw/deploy mini-batch      &        &  26.2 \\
    load $\pi$                 & 205    & 209 \\
    update $\phi$              &  74    &  74 \\
    \hline
  \end{tabular}
  \caption{The most important stages in the execution; com-Friendster on 65
  compute nodes, with 12K communities. Times in ms per iteration. The lower
  half of the table shows substages within \textit{update\_phi}. In the
  pipelining version, these components are done in parallel.}
  \label{table-pipeline}
\end{table}

@@

%################################################
\subsection{Horizontal vs. Vertical Scalability}
\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=\textwidth}
  \caption{Performance comparison between the distributed implementation
  running on DAS5 and the multi-threaded solution on machine with 40 cores and
  1TB of RAM.}
  \label{fig-scale-up}
\end{figure*}
One of the main drawbacks of designing a distributed solution for a given
algorithm is the inherent complexity of communication and synchronization. A
considerably simpler approach would be developing a multi-threaded version and
running it on a machine with abundant memory and CPU cores. In such a context,
access to all of the algorithm's state would be an order of magnitude faster
than RDMA. Additionally, the overhead associated with synchronizing threads is
negligible compared to using MPI primitives. To evaluate the efficacy of both
approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
memory. The physical machine underlying
the HPC Cloud system contains 40~CPU cores and does not oversubscribe
resources. Therefore, by provisioning all 40~cores we ensured that there is no
resource contention from other users of the system. Figure~\ref{fig-scale-up}
reports the execution time per algorithm iteration for two experimental setups.
%
First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
algorithm on the HPC Cloud
system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
test uses the com-DBLP data set to enable the use of a large number of
communities without running out of memory. It is clear from the results that
the performance can benefit from the additional cores provided by HPC Cloud system.
%
Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
system compared to 64 nodes of the DAS5
cluster using the com-Friendster data set. Clearly, the parallel and
distributed implementation vastly
outperforms the single-node multi-threaded solution. Moreover, the trajectory
of both curves shows a widening gap between them suggesting that the relative
performance difference will increase for larger $K$.
%
In conclusion, the overhead of network communication in the distributed version
is more than compensated by the increasing compute power compared to
a single-node implementation.

@@

\begin{figure}[b] % [htbp]
  \centering
  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
  \caption{Performance of DKV store compared to qperf.}
  \label{fig-qperf}
\end{figure}

@@

%################################################
\subsection{Convergence of Large Datasets}
\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx.eps, width=\textwidth}
  \caption{Convergence time of 6 different data sets.}
  \label{fig-ppx}
\end{figure*}

>>>>

@@

\begin{figure}[htb] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=0.48\textwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

In order to evaluate the horizontal scalability of the distributed
implementation we tested the system's performance across different cluster
sizes while holding the problem size constant. For this study, we used the
com-Friendster graph as it contains the largest number of vertices and edges.
%
Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
iterations across multiple cluster sizes. The x-axis starts from 8 worker nodes
as the data set is too large to fit into the collective memory of a smaller
cluster.
%
As shown in the figure, the execution
time steadily decreases by increasing the cluster size.
A deeper analysis of the individual execution phases of the algorithm provides
insights into the scalability of its building blocks.
%
In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
presents the cumulative time spent in individual computational phases across
iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
achieved for the same experiments reported in
Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
the execution is \textit{update\_phi\_pi}.
The reported total time for each cluster size is significantly less than the
sum of the execution times of the individual phases. This is due to the
overlapping execution of the two most expensive phases, namely,
\textit{update\_phi\_pi} and
mini-batch deployment. Both of these phases initially gain significant speedup with the
addition of compute nodes. However, the speedup curve gradually slows down for
larger cluster sizes as the work granularity of each worker node decreases,
limiting their resource utilization.
%
The time spent in \textit{update\_beta\_theta} remains relatively constant across cluster sizes
as it performs an insignificant amount of work compared to the synchronization
overhead of a collective MPI reduction operation contained within it.

@@

\begin{figure}[htb] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
  \caption{(a) Average execution time per algorithm iteration varying
  the number of latent communities proportionally to the number of compute
  nodes. (b) The exact number of communities used for each data point in (a).}
  \label{fig-weak-scaling}
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{l d{5.1} d{5.1}}
    \textbf{Iteration stages}  & \multicolumn{1}{c}{non-pipelined}
                                          & \multicolumn{1}{c}{pipelined} \\
    \hline
    total                      & 450    & 365 \\
    draw/deploy mini-batch      &  45.6  &      \\
    update\_phi                & 285    & 241 \\
    update\_pi                 &   3.8  &   4.6 \\
    update $\beta/\theta$      &  25.9  &   33.6 \\
    \hline
    \textbf{Substages within update\_phi} \\
    draw/deploy mini-batch      &        &  26.2 \\
    load $\pi$                 & 205    & 209 \\
    update $\phi$              &  74    &  74 \\
    \hline
  \end{tabular}
  \caption{The most important stages in the execution; com-Friendster on 65
  compute nodes, with 12K communities. Times in ms per iteration. The lower
  half of the table shows sub-stages within \textit{update\_phi}. In the
  pipelining version, these components are done in parallel.}
  \label{table-pipeline}
\end{table}

@@

%################################################
\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[htbp] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=0.48\textwidth}
  \caption{Performance comparison between the distributed implementation
  running on DAS5 and the multi-threaded solution on machine with 40 cores and
  1TB of RAM.}
  \label{fig-scale-up}
\end{figure}
One of the main drawbacks of designing a distributed solution for a given
algorithm is the inherent complexity of communication and synchronization. A
considerably simpler approach would be developing a multi-threaded version and
running it on a machine with abundant memory and CPU cores. In such a context,
access to all of the algorithm's state would be an order of magnitude faster
than RDMA. Additionally, the overhead associated with synchronizing threads is
negligible compared to using MPI primitives. To evaluate the efficacy of both
approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
memory. The physical machine underlying
the HPC Cloud system contains 40~CPU cores and does not oversubscribe
resources. Therefore, by provisioning all 40~cores we ensured that there is no
resource contention from other users of the system. Figure~\ref{fig-scale-up}
reports the execution time per algorithm iteration for two experimental setups.
%
First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
algorithm on the HPC Cloud
system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
test uses the com-DBLP data set to enable the use of a large number of
communities without running out of memory. It is clear from the results that
the performance can benefit from the additional cores provided by HPC Cloud system.
%
Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
system compared to 64 nodes of the DAS5
cluster using the com-Friendster data set. Clearly, the parallel and
distributed implementation vastly
outperforms the single-node multi-threaded solution. Moreover, the trajectory
of both curves shows a widening gap between them suggesting that the relative
performance difference will increase for larger $K$.
%
In conclusion, the overhead of network communication in the distributed version
is more than compensated by the increasing compute power compared to
a single-node implementation.

@@

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
  \caption{Performance of DKV store compared to qperf.}
  \label{fig-qperf}
\end{figure}

\begin{figure*}[htbp] % [htb]
  \centering
  \epsfig{file=plots/ppx.eps, width=\textwidth}
  \caption{Convergence time of 6 different data sets.}
  \label{fig-ppx}
\end{figure*}

@@

%################################################
\subsection{Convergence of Large Datasets}
